---
title: "Untitled"
author: "Kinjal"
date: "2 October 2017"
output: html_document
---

---
title: "DAND_STATS"
author: "Kinjal"
date: "1 October 2017"
output: html_document
---

#Title : Analysis of Stroops file

About Stroop's effect ?

In a Stroop task, participants are presented with a list of words, with each word displayed in a color of ink. The participant's task is to say out loud the color of the ink in which the word is printed. The task has two conditions: a congruent words condition, and an incongruent words condition. In the congruent words condition, the words being displayed are color words whose names match the colors in which they are printed: for example RED, BLUE. In the incongruent words condition, the words displayed are color words whose names do not match the colors in which they are printed: for example PURPLE, ORANGE. In each case, we measure the time it takes to name the ink colors in equally-sized lists. Each participant will go through and record a time from each condition.


1) Identify the independent and dependent variables in the experiment.

Understanding types of variables :
Independent variables are experimental or predictor variables whereas dependent variables are final outcomes.

Based upon above explanation 

Independent variables are :
Word Condition of congruency/incrogruency.

Dependent Variables is :
Time in seconds taken to read out the words correctly.

2) What is the Null and Alternate hypothesis for the experiment.

Null hypothesis : Time in seconds taken to read out incongruent words will be less or same than that of congruent words.Hence there will be no effect of incongruency.
H0: mui <= muc (mui - population mean of incongruent values, muc - population mean of congruent values)

Alternate hypotesis : Time in seconds taken to read out incongruent words will be more of than that of congruent words.Hence there will be effect of incongruency.
HA: mui > muc (mui - population mean of incongruent values, muc - population mean of congruent values)

3) Statistical test used.

I have used a t-test over z test because population parameters are not know in our dataset and sample size is also small.
I have decided to choose a one tailed test is right direction, because my alternate hypothesis will have incongruent reading time more than congruent reading time.Hence I have I want to analyse my direction of test.
The test will be repeated order dependent sample test because the test is within subject design , where each subject is assinged 
two conditions in random order.
I have choosen 95% of confidence interval than using 98% confidence interval because looking at the size of the dataset , we are 
satisfied with lower chance of creating confidence , whereas a 98% confidence interval would have required greater sample size.



### Reading the csv file 
```{r}
stroop <- read.csv("C:\\Users\\KACHI\\Desktop\\stroopdata.csv")

```


4) Descriptive statistics, including at least one measure of centrality and one measure of variability, have been computed for the dataset's groups.

Below the mean , median , standard deviation(sd) and variance has been created.

```{r}
# Tidy up the data for later analysis
library(tidyr); suppressMessages(library(dplyr)) 
# Add a column identifying subjects
stroop.subject <- mutate(stroop, subject = 1:nrow(stroop))
# Tidy up data by keeping one variable in one column
tidy.stroop <- gather(stroop.subject, congruency, time, -subject)
# Calculate the average time for both groups
tidy.stroop %>%
    group_by(congruency) %>%
    summarise(mean(time), median(time), sd(time), var(time))
```


5) One or two visualizations have been created that show off the data, including comments on what can be observed in the plot or plots.


```{r }
stroop$diff <- stroop$Incongruent - stroop$Congruent

```

Diff is a variable that has difference values of congruent and incongruent columns.

### Below is the ggplot of difference along with mean , median amd 10 % to 90 % interval.

```{r  message=FALSE, warning=FALSE}

library(ggplot2)
library(gridExtra)

q1 <- ggplot(aes(x=diff),
       data =  stroop)+
   geom_histogram(color =I('black'),
                  fill = I('#099009'))+
   geom_vline(xintercept = mean(stroop$diff),
              col = "royalblue")+
   geom_vline(xintercept = median(stroop$diff),
              col = "red")+
   geom_vline(xintercept = quantile(probs = 0.9 ,
                                    stroop$diff), 
              linetype = 2)+
   geom_vline(xintercept = quantile(probs = 0.1 ,
                                    stroop$diff),
              linetype = 2)

q2 <- ggplot(aes(x= Congruent),
       data =  stroop)+
   geom_histogram(color =I('black'),
                  fill = I('#099009'))

q3 <- ggplot(aes(x= Incongruent),
       data =  stroop)+
   geom_histogram(color =I('black'),
                  fill = I('#099009'))

grid.arrange(q3, q2,q1, ncol =1)

```
There are three plots of Incongruence , Congruence and the difference between them.
As it can be seen the the mean and median of the differences is around 7.5 , there is a drop between 12.5 to 16.5 and between 17.5 to 18.5.



6) A statistical test has been correctly performed and reported, including test statistic, p-value, and test result. The test results are interpreted in terms of the experimental task performed.

## Below I have calculated t statistics using a simple in built function .
References for following t test :
http://www.instantr.com/2012/12/29/performing-a-one-sample-t-test-in-r/


```{r}
# Verify using the t.test() function
t.test(x=stroop$Incongruent, y=stroop$Congruent,
       alternative = "greater", 
       paired = TRUE, 
       conf.level = 0.95)
```

From the following information :
n = 24
DOF : n-1 = 23
alpha = 0.05
t critical value obtained from t table is 2.05 , hence result is statistically important so we have to reject the null.

